---
title: "Task1"
author: "Jin Chen"
date: "9/13/2020"
output: html_document
---

```{r Load libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(skimr)
library(ggthemes)
library(viridis)
library(RColorBrewer)

library(data.table)
library(ggmosaic)
```

```{r Load data}
QVI_behaviour <- read_csv("../data/QVI_purchase_behaviour.csv")
QVI_transact <- read_xlsx("../data/QVI_transaction_data.xlsx")
```

```{r Exploratory}
str(QVI_transact)
```

```{r Convert DATE to correct format}
QVI_transact <- QVI_transact %>% 
  mutate(DATE = as.Date(DATE, origin = "1899-12-30")) 

```

```{r Summary of PROD_NAME}
summary(QVI_transact$PROD_NAME)
QVI_transact %>% 
  select(PROD_NAME) %>% 
  head(10)


```

```{r Further examine PROD_NAME}
#### Examine the words in PROD_NAME to see if there are any incorrect entries
#### such as products that are not chips
productWords <- data.table(unlist(strsplit(unique(QVI_transact$PROD_NAME), "
")))
setnames(productWords, 'words')
```

```{r}
#### Removing digits
productWords <- productWords %>% 
  mutate(words = str_replace_all(words, "[:digit:]", " "))

#### Removing special characters
productWords <- productWords %>% 
  mutate(words = str_replace_all(words, "[[:punct:]]", " "))

productWords <- productWords %>% 
  mutate(words = str_replace_all(words, " [gG]", " ")) #replace the g and G for grams, but only if there is a space before it

#### Let's look at the most common words by counting the number of times a word appears and 
#### sorting them by this frequency in order of highest to lowest frequency

sepWords <- productWords %>% 
  select(words) %>% 
  str_split(" ")

freqWords <- table(unlist(sepWords))
freqWords <- as.data.frame(freqWords)
freqWords %>% 
  arrange(desc(Freq))

```
There are salsa products in the dataset but we are only interested in the chips
category, so let's remove these.

```{r}
#### Remove salsa products
QVI_transact <- QVI_transact %>% 
  filter(!str_detect(tolower(PROD_NAME), "salsa"))

# check that salsa products are removed
QVI_transact %>% 
  filter(str_detect(tolower(PROD_NAME), "salsa"))

```
Next, we can use `summary()` to check summary statistics such as mean, min and max
values for each feature to see if there are any obvious outliers in the data and if
there are any nulls in any of the columns (`NA's : number of nulls` will appear in
the output if there are any nulls).

```{r initial summary}
#### Summarise the data to check for nulls and possible outliers
# Over to you!

summary(QVI_transact)
```

There are no nulls in the columns but product quantity appears to have an outlier
which we should investigate further. Let's investigate further the case where 200
packets of chips are bought in one transaction.
```{r }
#### Filter the dataset to find the outlier
# Over to you! Use a filter to examine the transactions in question.
QVI_transact %>% 
  filter(PROD_QTY == 200)

#### Let's see if the customer has had other transactions
# Over to you! Use a filter to see what other transactions that customer made.
QVI_transact %>% 
  filter(LYLTY_CARD_NBR == 226000)
```
It looks like this customer has only had the two transactions over the year and is
not an ordinary retail customer. The customer might be buying chips for commercial
purposes instead. We'll remove this loyalty card number from further analysis.

```{r}
#### Filter out the customer based on the loyalty card number
# Over to you!
QVI_transact <- QVI_transact %>% 
  filter(LYLTY_CARD_NBR != 226000)

#### Re-examine transaction data
# Over to you!
summary(QVI_transact)
```

That's better. Now, let's look at the number of transaction lines over time to see
if there are any obvious data issues such as missing data.
```{r}
#### Count the number of transactions by date
# Over to you! Create a summary of transaction count by date.
QVI_transact %>% 
  group_by(DATE) %>% 
  summarize(count = n())

```
There's only 364 rows, meaning only 364 dates which indicates a missing date. Let's
create a sequence of dates from 1 Jul 2018 to 30 Jun 2019 and use this to create a
chart of number of transactions over time to find the missing date.

```{r fig.align = "center"}
#### Create a sequence of dates and join this the count of transactions by date
# Over to you - create a column of dates that includes every day from 1 Jul 2018 to 30 Jun 2019, and join it onto the data to fill in the missing day.
seqDates <- data.frame("DATE" = seq(from = as.Date("2018-07-01"), to = as.Date("2019-06-30"), by = 'day'))

transactions_by_day <- QVI_transact %>% 
  group_by(DATE) %>% 
  summarize(count = n()) %>% 
  right_join(seqDates) %>% 
  arrange(DATE)
transactions_by_day 

#set NA to 0
transactions_by_day <- transactions_by_day %>% 
  mutate(count = replace_na(count, 0))

#### Setting plot themes to format graphs
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
#### Plot transactions over time
ggplot(transactions_by_day, aes(x = DATE, y = count)) +
 geom_line() +
 labs(x = "Day", y = "Number of transactions", title = "Transactions over time") +
 scale_x_date(breaks = "1 month") +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```
We can see that there is an increase in purchases in December and a break in late
December. Let's zoom in on this.
```{r fig.align = "center"}
#### Filter to December and look at individual days
# Over to you - recreate the chart above zoomed in to the relevant dates.

transactions_december18 <- transactions_by_day %>% 
  filter(DATE >= "2018-12-01" & DATE <= "2018-12-31")

ggplot(transactions_december18, aes(x = DATE, y = count)) +
 geom_line() +
 labs(x = " ", y = " ", title = "Transactions In December 2018") +
 scale_x_date(breaks = "day") +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

We can see that the increase in sales occurs in the lead-up to Christmas and that
there are zero sales on Christmas day itself. This is due to shops being closed on
Christmas day.

Now that we are satisfied that the data no longer has outliers, we can move on to
creating other features such as brand of chips or pack size from PROD_NAME. We will
start with pack size.

```{r Create pack size}
#### Pack size
#### We can work this out by taking the digits that are in PROD_NAME

QVI_transact <- QVI_transact %>% 
  mutate(PACK_SIZE = as.integer(str_extract(PROD_NAME, "\\d+")))


#### Always check your output
#### Let's check if the pack sizes look sensible
QVI_transact %>% 
  select(PACK_SIZE) %>% 
  arrange(desc(PACK_SIZE))

```
The largest size is 380g and the smallest size is 70g - seems sensible!
```{r }
#### Let's plot a histogram of PACK_SIZE since we know that it is a categorical variable and not a continuous variable even though it is numeric.
# Over to you! Plot a histogram showing the number of transactions by pack size.

QVI_transact %>% 
  group_by(PACK_SIZE) %>% 
  ggplot(aes(x = PACK_SIZE)) +
  geom_histogram()


```
Pack sizes created look reasonable.
Now to create brands, we can use the first word in PROD_NAME to work out the brand
name

```{r Create brand name}
#### Brands
# Over to you! Create a column which contains the brand of the product, by extracting it from the product name.
QVI_transact <- QVI_transact %>% 
  mutate(Brand = word(QVI_transact$PROD_NAME, 1, 1)) 


#### Checking brands
# Over to you! Check the results look reasonable.
QVI_transact %>% 
  select(Brand) %>% 
  count(Brand, sort = TRUE) %>% 
  head(10) 
```
Some of the brand names look like they are of the same brands - such as RED and
RRD, which are both Red Rock Deli chips. Let's combine these together.

```{r Clean brand names}
#### Clean brand names
# Over to you! Add any additional brand adjustments you think may be required.
QVI_transact <- QVI_transact %>% 
  mutate(Brand = case_when(Brand == "Red" ~ "RRD",
                           Brand == "SNBTS" ~ "Sunbites",
                           Brand == "Snbts" ~ "Sunbites",
                           Brand == "INFZNS" ~ "Infuzions",
                           Brand == "Infzns" ~ "Infuzions",
                           Brand == "WW" ~ "Woolworths",
                           Brand == "WOOLWORTHS" ~ "Woolworths",
                           Brand == "SMITH" ~ "Smiths",
                           Brand == "Smith" ~ "Smiths",
                           Brand == "NCC" ~ "Natural",
                           Brand == "NATURAL" ~ "Natural",
                           Brand == "Dorito" ~ "Doritos",
                           Brand == "Grain" ~ "GrnWves",
                           Brand == "GRNWVES" ~ "GrnWves",
                           TRUE ~ Brand))
#### Check again
# Over to you! Check the results look reasonable.
QVI_transact %>% 
  distinct(Brand) %>% 
  arrange(Brand)
```

### Examining customer data
Now that we are happy with the transaction dataset, let's have a look at the
customer dataset.

```{r 1 Exploratory data analysis}
#### Examining customer data
# Over to you! Do some basic summaries of the dataset, including distributions of any key columns.
skim(QVI_transact)

```

```{r }
#### Merge transaction data to customer data
data <- merge(QVI_transact, QVI_behaviour, all.x = TRUE)
```
As the number of rows in `data` is the same as that of `transactionData`, we can be
sure that no duplicates were created. This is because we created `data` by setting
`all.x = TRUE` (in other words, a left join) which means take all the rows in
`transactionData` and find rows with matching values in shared columns and then
joining the details in these rows to the `x` or the first mentioned table.

Let's also check if some customers were not matched on by checking for nulls.
```{r Check for missing customer details}
# Over to you! See if any transactions did not have a matched customer.
data %>% 
  filter_all(any_vars(is.na(.))) 

#apply(data, 2, function(x) any(is.na(x)))

```
Great, there are no nulls! So all our customers in the transaction data has been
accounted for in the customer dataset.
Note that if you are continuing with Task 2, you may want to retain this dataset
which you can write out as a csv

```{r Code to save dataset as a csv}
#fwrite(data, paste0(filePath,"QVI_data.csv"))
write_csv(data, path = "QVI_data.csv")

```

Data exploration is now complete!
## Data analysis on customer segments
Now that the data is ready for analysis, we can define some metrics of interest to the client:
Who spends the most on chips (total sales), describing customers by lifestage and how premium their general purchasing behaviour is
- How many customers are in each segment
- How many chips are bought per customer by segment
- What's the average chip price by customer segment

We could also ask our data team for more information. Examples are:
- The customer's total spend over the period and total spend for each transaction to understand what proportion of their grocery spend is on chips
- Proportion of customers in each customer segment overall to compare against the mix of customers who purchase chips

Let's start with calculating total sales by LIFESTAGE and PREMIUM_CUSTOMER and
plotting the split by these segments to describe which customer segment contribute most to chip sales.

```{r fig.width = 20, fig.height = 5}
#### Total sales by LIFESTAGE and PREMIUM_CUSTOMER
# Over to you! Calculate the summary of sales by those dimensions and create a plot.

Total_sales <- data %>% 
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>% 
  summarize(sales = sum(TOT_SALES)) %>% 
  arrange(desc(sales))

Total_sales %>% 
  ggplot(aes(x = tidytext::reorder_within(LIFESTAGE, sales, PREMIUM_CUSTOMER), y = sales, fill = PREMIUM_CUSTOMER)) +
  geom_histogram(stat = "identity", show.legend = FALSE) +
  facet_wrap(~PREMIUM_CUSTOMER, scales = "free_y") +
  coord_flip() +
  tidytext::scale_x_reordered() +
  labs(x = "", y = "Sales in Dollars", title = "Sales vary by customer type")
```

There are more Mainstream - young singles/couples and Mainstream - retirees who buy chips. This contributes to there being more sales to these customer segments but this is not a major driver for the Budget - Older families segment. (? Budget - Older families segment has a large sales amount)

Higher sales may also be driven by more units of chips being bought per customer.

Let's have a look at this next.

```{r fig.width = 10, fig.align = "center"}
#### Average number of units per customer by LIFESTAGE and PREMIUM_CUSTOMER
# Over to you! Calculate and plot the average number of units per customer by those two dimensions.

## units per customer is sum of PROD_QTY divided by unique number of card numbers per customer segment
data %>% 
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>% 
  summarize(Units = sum(PROD_QTY)/n_distinct(LYLTY_CARD_NBR)) %>% 
  ggplot(aes(x = LIFESTAGE, y = Units, fill = PREMIUM_CUSTOMER)) +
  geom_histogram(stat = "identity", position = "dodge") +
  labs(x = " ", y = " ", title = "Avg Units per Transaction by Customer Segment") +
  theme_minimal() +
  coord_flip() +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_viridis(discrete = TRUE) 

```
Older families and young families in general buy more chips per customer
Let's also investigate the average price per unit chips bought for each customer
segment as this is also a driver of total sales.

```{r fig.width = 10, fig.align = "center"}
#### Average price per unit by LIFESTAGE and PREMIUM_CUSTOMER
# Over to you! Calculate and plot the average price per unit sold (average sale price) by those two customer dimensions.

## Price per unit is sum of total sales divided by sum of total product quantity
data %>% 
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>% 
  summarize(priceperUnits = sum(TOT_SALES)/sum(PROD_QTY)) %>% 
  ggplot(aes(x = LIFESTAGE, y = priceperUnits, fill = PREMIUM_CUSTOMER)) +
  geom_histogram(stat = "identity", position = "dodge") +
  labs(x = " ", y = " ", title = "Avg Price Per Unit Sold by Customer Segment") +
  theme_minimal() +
  coord_flip(ylim=c(3, 4)) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_viridis(discrete = TRUE)



```
Mainstream midage and young singles and couples are more willing to pay more per
packet of chips compared to their budget and premium counterparts. This may be due
to premium shoppers being more likely to buy healthy snacks and when they buy
chips, this is mainly for entertainment purposes rather than their own consumption.

This is also supported by there being fewer premium midage and young singles and
couples buying chips compared to their mainstream counterparts.
As the difference in average price per unit isn't large, we can check if this difference is statistically different.

```{r}
#### Perform an independent t-test between (mainstream) vs (premium and budget) (midage) and (young singles and couples)
# Over to you! Perform a t-test to see if the difference is significant.

#is there an effect between mainstream midage & youngsingles/couples and premium & budget midage & youngsingles/couples on price per unit?

price <- data %>% 
  group_by(LIFESTAGE, PREMIUM_CUSTOMER) %>% 
  summarize(avgpriceperUnit = sum(TOT_SALES)/sum(PROD_QTY)) %>% 
  mutate(Label = case_when(PREMIUM_CUSTOMER == "Mainstream" ~ "Mainstream",
                           PREMIUM_CUSTOMER != "Mainstream" ~ "Nonmainstream")) %>% 
  filter(LIFESTAGE %in% c("YOUNG SINGLES/COUPLES", "MIDAGE SINGLES/COUPLES"))


# priceperTransaction <- data %>% 
#   mutate(Price = TOT_SALES/PROD_QTY,
#          Label = case_when(PREMIUM_CUSTOMER == "Mainstream" ~ "Mainstream",
#                            PREMIUM_CUSTOMER != "Mainstream" ~ "Nonmainstream")) %>% 
#   select(LIFESTAGE, PREMIUM_CUSTOMER,TOT_SALES, PROD_QTY, Price, Label) %>% 
#   filter(LIFESTAGE %in% c("YOUNG SINGLES/COUPLES", "MIDAGE SINGLES/COUPLES"))
# priceperTransaction

infer::t_test(x = price,
              formula = avgpriceperUnit ~ Label,
              order = c("Mainstream", "Nonmainstream"))
 

```

The t-test results in a p-value of *0.031128* (which is less than 0.05), i.e. the unit price for mainstream, young and mid-age singles and couples *ARE* significantly higher than
that of budget or premium, young and midage singles and couples.


## Deep dive into specific customer segments for insights
We have found quite a few interesting insights that we can dive deeper into.
We might want to target customer segments that contribute the most to sales to
retain them or further increase sales. Let's look at Mainstream - young
singles/couples. For instance, let's find out if they tend to buy a particular
brand of chips.

We might want to target customer segments that contribute the most to sales to
retain them or further increase sales. Let's look at Mainstream - young
singles/couples. For instance, let's find out if they tend to buy a particular
brand of chips.


```{r fig.align = "center"}
#### Deep dive into Mainstream, young singles/couples
# Over to you! Work out if there are brands that these two customer segments prefer more than others. You could use a technique called affinity analysis or a-priori analysis (or any other method if you prefer)

data %>% 
  group_by(Brand) %>% 
  ggplot(aes(x = Brand)) +
  geom_bar() +
  coord_flip()

 data %>% 
  group_by(LIFESTAGE, PREMIUM_CUSTOMER, Brand) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = Brand, y = count, fill = LIFESTAGE)) +
  geom_histogram(stat = "identity", position = "dodge") +
  facet_wrap(~PREMIUM_CUSTOMER) +
  coord_flip()

 brand <- data %>% 
  group_by(LIFESTAGE, PREMIUM_CUSTOMER, Brand) %>% 
  summarize(count = n()) %>% 
  filter(LIFESTAGE == "YOUNG SINGLES/COUPLES",
         PREMIUM_CUSTOMER == "Mainstream")
 
 
 
segment1 <- data %>% 
  filter(PREMIUM_CUSTOMER == "Mainstream" & LIFESTAGE == "YOUNG SINGLES/COUPLES")

segment2 <- data %>% 
  filter(!(PREMIUM_CUSTOMER == "Mainstream" & LIFESTAGE == "YOUNG SINGLES/COUPLES"))

segment1 %>% 
  group_by(Brand) %>% 
  summarize(count = n()) %>% 
  mutate(prop = count/sum(count)) %>% 
  arrange(desc(prop)) %>% 
  ggplot(aes(x = fct_reorder(Brand, prop), y = prop)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Top Brands Bought by Mainstream - Young Singles/Couples",
       x = " " )
  

```
We can see that :
The Mainstream - Yung Singles/Couples customer segment primarily buy Kettle, Doritos and Pringles with Smiths following closely behind.

Let's also find out if our target segment tends to buy larger packs of chips.

```{r fig.align = "center"}
#### Preferred pack size compared to the rest of the population
# Over to you! Do the same for pack size.

segment1 %>% 
  group_by(PACK_SIZE) %>% 
  summarize(count = n()) %>% 
  mutate(prop = count/sum(count),
         PACK_SIZE = as.factor(as.integer(PACK_SIZE))) %>% 
  arrange(desc(PACK_SIZE)) %>% 
  ggplot(aes(x = fct_reorder(PACK_SIZE, prop, .desc = TRUE), y = prop)) +
  geom_histogram(stat = "identity", show.legend = FALSE) +
  theme_minimal() +
  labs(title = "Top Sizes Bought by Mainstream - Young Singles/Couples",
       x = " ",
       y = "Proportion of Sales")
  
segment2 %>% 
  group_by(PACK_SIZE) %>% 
  summarize(count = n()) %>% 
  mutate(prop = count/sum(count),
         PACK_SIZE = as.factor(as.integer(PACK_SIZE))) %>% 
  arrange(desc(PACK_SIZE)) %>% 
  ggplot(aes(x = fct_reorder(PACK_SIZE, prop, .desc = TRUE), y = prop)) +
  geom_histogram(stat = "identity", show.legend = FALSE) +
  theme_minimal() +
  labs(title = "Top Sizes Bought by Other Segments",
       x = " ",
       y = "Proportion of Sales")

```
The top preferred pack sizes for both customer segments are 175g, 150g, and 134g

